---
title: "US Border Apprehension Analysis"
author: "Udochi Maduakor,Ozair Meghani,Junpei Xiao"
date: "February 17, 2018"
output: ioslides_presentation
css: style.css
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(dplyr)
library(DT)
library(tidyverse)
library(knitr)
library(kableExtra)

```
## **INTRODUCTION** {.flexbox .vcenter}

<div class="centered">

Illegal entry to the United States has been of great concern to the US customs and border protection unit. Our analysis below will clarify if the recent enforcement of the immigration Laws by president had a positive effect on the border apprenshions between the two years. </div>


## **Contents**{.flexbox .vcenter}

- Review of Apprehensions in 2010
- Review of Apprehensions in 2017
- Graphical comparison of border apprehensions by sector
- Graphical comparison of border apprehensions by Month
- T-Test Analysis
- T-Test Analysis Continued


##Review of Apprehensions 2010
```{r}
A2010 <- read.csv("BP Apprehensions 2010.csv", header = TRUE, stringsAsFactors = FALSE)
rownames(A2010) <- A2010[,1]
A2010 <-  subset(A2010, select= -c(Sector))
A2010 <- rbind(A2010, colSums(A2010))
rownames(A2010) <- c(rownames(A2010)[-length(rownames(A2010))], "Total")
A2010 <- cbind(A2010,rowSums(A2010))
colnames(A2010) <- c(colnames(A2010)[-length(colnames(A2010))], "Total")
barplot(A2010[1:9,13], names.arg = rownames(A2010)[1:9], 
        las=2,
        axisnames=TRUE,
        main="2010 Border Patrol Apprehensions by Sector",
        border="blue",
        col="yellow")
```

##Review of Apprehensions 2017
```{r}
A2017 <- read.csv("BP Apprehensions 2017.csv", header = TRUE, stringsAsFactors = FALSE)
rownames(A2017) <- A2017[,1]
A2017 <-  subset(A2017, select= -c(Sector))
A2017 <- rbind(A2017, colSums(A2017))
rownames(A2017) <- c(rownames(A2017)[-length(rownames(A2017))], "Total")
A2017 <- cbind(A2017,rowSums(A2017))
colnames(A2017) <- c(colnames(A2017)[-length(colnames(A2017))], "Total")
barplot(A2017[1:9,13], names.arg = rownames(A2010)[1:9], 
        las=2,
        axisnames=TRUE,
        main="2017 Border Patrol Apprehensions by Sector",
        border="blue",
        col="RED")
```

##**Graphical comparison of border apprehensions by sector**
```{r}
A2010 <- read.csv("BP Apprehensions 2010.csv", header = TRUE, stringsAsFactors = FALSE)
A2017 <- read.csv("BP Apprehensions 2017.csv", header = TRUE, stringsAsFactors = FALSE)
rownames(A2010) <- A2010[,1]
A2010 <-  subset(A2010, select= -c(Sector))
## rbind ColSums to dataframe
A2010 <- rbind(A2010, colSums(A2010))
## rename the row with column totals "Total"
rownames(A2010) <- c(rownames(A2010)[-length(rownames(A2010))], "Total")
## cbind rowSums to dataframd
A2010 <- cbind(A2010,rowSums(A2010))
## rename last column "Totals
colnames(A2010) <- c(colnames(A2010)[-length(colnames(A2010))], "Total")

rownames(A2017) <- A2017[,1]
A2017 <-  subset(A2017, select= -c(Sector))
## rbind ColSums to dataframe
A2017 <- rbind(A2017, colSums(A2017))
## rename the row with column totals "Total"
rownames(A2017) <- c(rownames(A2017)[-length(rownames(A2017))], "Total")
## cbind rowSums to dataframd
A2017 <- cbind(A2017,rowSums(A2017))
## rename last column "Totals
colnames(A2017) <- c(colnames(A2017)[-length(colnames(A2017))], "Total")

compsect2010 <- as.data.frame(matrix(c(A2010[1,13],A2010[2,13],A2010[3,13],A2010[4,13],A2010[5,13],A2010[6,13],A2010[7,13],A2010[8,13],A2010[9,13]), nrow = 1))
colnames(compsect2010) <- rownames(A2010)[1:9]
compsect2017 <- as.data.frame(matrix(c(A2017[1,13],A2017[2,13],A2017[3,13],A2017[4,13],A2017[5,13],A2017[6,13],A2017[7,13],A2017[8,13],A2017[9,13]), nrow = 1))
colnames(compsect2017) <- rownames(A2010)[1:9]
compsect10.17 <- rbind(compsect2010, compsect2017)
barplot(as.matrix(compsect10.17), beside = TRUE, col = c("red", "blue"), bty="n", ylab = "Apprehensions", xlab = "Sector", las=2, cex.lab = 1.5, cex.axis = 0.7, cex.names = 0.6)
legend("topleft", c("2010","2017"), pch=15,  col=c("red","blue"),  bty="n")        
```


##**Graphical comparison of border apprehensions by Month**
```{r}
A2010 <- read.csv("BP Apprehensions 2010.csv", header = TRUE, stringsAsFactors = FALSE)
A2017 <- read.csv("BP Apprehensions 2017.csv", header = TRUE, stringsAsFactors = FALSE)
rownames(A2010) <- A2010[,1]
A2010 <-  subset(A2010, select= -c(Sector))
## rbind ColSums to dataframe
A2010 <- rbind(A2010, colSums(A2010))
## rename the row with column totals "Total"
rownames(A2010) <- c(rownames(A2010)[-length(rownames(A2010))], "Total")
## cbind rowSums to dataframd
A2010 <- cbind(A2010,rowSums(A2010))
## rename last column "Totals
colnames(A2010) <- c(colnames(A2010)[-length(colnames(A2010))], "Total")

rownames(A2017) <- A2017[,1]
A2017 <-  subset(A2017, select= -c(Sector))
## rbind ColSums to dataframe
A2017 <- rbind(A2017, colSums(A2017))
## rename the row with column totals "Total"
rownames(A2017) <- c(rownames(A2017)[-length(rownames(A2017))], "Total")
## cbind rowSums to dataframd
A2017 <- cbind(A2017,rowSums(A2017))
## rename last column "Totals
colnames(A2017) <- c(colnames(A2017)[-length(colnames(A2017))], "Total")

compmnth2010 <- as.data.frame(matrix(c(A2010[10,1],A2010[10,2],A2010[10,3],A2010[10,4],A2010[10,5],A2010[10,6],A2010[10,7],A2010[10,8],A2010[10,9],A2010[10,10],A2010[10,11],A2010[10,12]), nrow = 1))
colnames(compmnth2010) <- colnames(A2010)[1:12]
compmnth2017 <- as.data.frame(matrix(c(A2017[10,1],A2017[10,2],A2017[10,3],A2017[10,4],A2017[10,5],A2017[10,6],A2017[10,7],A2017[10,8],A2017[10,9],A2017[10,10],A2017[10,11],A2017[10,12]), nrow = 1))
colnames(compmnth2017) <- colnames(A2010)[1:12]
compmnth10.17 <- rbind(compmnth2010, compmnth2017)
barplot(as.matrix(compmnth10.17), beside = TRUE, col = c("red", "blue"), bty="n", ylab = "Apprehensions", xlab = "Month",las=2, cex.lab = 1.5, cex.axis = 0.7, cex.names= 0.7)
legend("topleft", c("2010","2017"), pch=15,  col=c("red","blue"),  bty="n")
```

##**T-Test Analysis**{.smaller}

### Max Apprehensions by Sector between 2010 and 2017

```{r}
maxsect2010 <- as.numeric(as.vector(A2010[8,1:12]))
maxsect2017 <- as.numeric(as.vector(A2017[6,1:12]))
t.test(maxsect2010, maxsect2017, var.equal = TRUE)
```

The results show that there is not a significant difference in means between the two sectors(Tucson 2010, Rio Grande Valley 2017) with Max apprehensions in 2010 and 2017

##**T-Test Analysis Continued** {.smaller}

### Max Apprehensions by 3-Month Period between 2010 and 2017

```{r}
max3mnth2010 <- c(A2010$March[1:9],A2010$April[1:9],A2010$May[1:9])
max3mnth2017 <- c(A2017$October[1:9], A2017$November[1:9], A2017$December[1:9])
t.test(max3mnth2010, max3mnth2017, var.equal = TRUE) 
```

The 3-month period in 2010 with max apprehensions was March, April, and May. The 3-month period in 2017 with max apprehensions was October, November, December. Based on the t-test, the results show that there is no significant difference in means between these two 3-month periods.  